An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations

### 2020-07-28 Notes/algorithm.tex
real time improved
new results + paper

\paragraph{Dynamical Model and Forecasting} 
Establishing a dynamical model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\[
\bY = f(\bX)\ .
\]
In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values. For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$. We will see, in section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us. That is why we consider here a naive dynamical model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\label{eq:dyn.model}
\end{equation}
where $\bA\in\RR^{M\times M}$. In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization} of a nonlinear phenomenon, when the system is sufficiently regular. Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}

@@

\paragraph{Signal Extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the signal has to be extended, the left ``side'' being fixed since it only concerns the past values of the signal. We therefore construct the extended signal $\tilde\bx\in\RR^{N+L}$ concatenating the observed signal $\bx$, and the forward forecasting $\tilde\bx_{\mathrm{fw}}\in\RR^L$. We summarize the extension step in Algorithm~\ref{alg:extension}. Notice that we could, if necessary, handle the backward estimation using the same strategy as described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.

@@

\subsection{Step 2: Extended Time-Frequency Representation}

Let $\ccF:\RR^{N}\to\CC^{F\times N}$ generically denotes the TF representation of interest to us, which could be, for instance, STFT, CWT, SST, or RS. Here, $F$ typically denotes the size of the discretization along the frequency axis. Due to the boundary effects, the representation $\ccF(\bx_N)$ shows undesired patterns when approaching its edges. To alleviate the boundary effects, we apply the representation to the estimated extended signal $\tilde\bx$. This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$. Finally, the boundary-effects insensitive representation $\ccF^{\mathrm{ext}}:\RR^{N}\to\CC^{F\times N}$ of $\bx_N$ is given for all $\nu\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
\begin{equation}
\ccF^{\mathrm{ext}}(\bx_N)[\nu,n] = \ccF(\tilde \bx_{N+L})[\nu,n]\ .
\label{eq:restriction}
\end{equation}
This amounts to restricting the representation $\ccF(\tilde \bx_{N+L})$ to the current measurement interval of $\bx_N$. For the sake of simplicity, we denote the restriction operator by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$. Consequently, we have:
\begin{equation}
\ccF^{\mathrm{ext}}(\bx_N) = \cR\left( \ccF(\tilde \bx_{N+L}) \right) \ .
\label{eq:bound.free.TFR}
\end{equation}
We call $\ccF^{\mathrm{ext}}$ the {\em boundary-free TF representation}.

@@

\begin{remark}
Unlike other representations, RS is not causal and rigorously requires knowing the STFT on the whole time-frequency plane before making any reassignment. Nevertheless, in~\cite{Lin17conceft}, the authors propose a causal RS that only requires the future values of the STFT within a limited time horizon (typically less than $L$). We use this version in the paper to enable real-time implementation of RS.
\end{remark}

>>>>

\paragraph{Dynamical Model and Forecasting} 
Establishing a dynamical model consists in determining the relation linking $\bY$ to $\bX$, that is finding a function $f$ so that
\[
\bY = f(\bX)\ .
\]
In a general framework, forecasting means estimating the function $f$ from the observed values taken by the signal, in order to predict its future values. For instance, the dynamic mode decomposition~\cite{Schmid10dynamic,Williams15data} or other more complicated models \cite{Roberts13Gaussian,DeLivera11forecasting,west2006bayesian,vlachas2018data}, allow this by setting additional constraints on the behavior of $f$. We will see, in section~\ref{se:theoretical}, that considering such a complex dynamic model is not necessary for the study of the oscillatory signals of interest to us. That is why we consider here a naive dynamical model, assuming that we have the following relation:
\begin{equation}
\bY = \bA\bX\ ,
\label{eq:dyn.model}
\end{equation}
where $\bA\in\RR^{M\times M}$. In other words, we adopt a classical strategy in the study of dynamical systems, the {\em linearization of a nonlinear system}, when the system is sufficiently regular. Notice that this linearized dynamical model can be written equivalently in function of the subsignals $\bx_k$, as:
\begin{equation}
\bx_{k+1} = \bA\bx_{k},\ ,\forall k\in\{0,\dots,K-1\}\ .
\end{equation}

@@

\paragraph{Signal Extension} Since we are building a real-time algorithm, we consider that only the right ``side'' of the signal has to be extended, the left ``side'' being fixed since it only concerns the past values of the signal. We therefore construct the extended signal $\tilde\bx\in\RR^{N+L}$ concatenating the observed signal $\bx$, and the forward forecasting $\tilde\bx_{\mathrm{fw}}\in\RR^L$. We summarize the extension step in Algorithm~\ref{alg:extension}. %We mention that we could, if necessary, handle the backward estimation using the same strategy as described above, but applying it to the reverse signal $\bx^{\rm r} = \begin{pmatrix} \bx[N-1] & \cdots & \bx[0] \end{pmatrix}^T$.

@@

\subsection{Step 2: Extended Time-Frequency Representation}
\label{sse:extension.TFR}

Let $\ccF:\RR^{N}\to\CC^{F\times N}$ generically denotes the TF representation of interest to us, which could be, for instance, STFT, CWT, SST, or RS. Here, $F$ typically denotes the size of the discretization along the frequency axis. Due to the boundary effects, the representation $\ccF(\bx_N)$ shows undesired patterns near its edges. To alleviate the boundary effects, we apply the representation to the estimated extended signal $\tilde\bx$. This strategy moves the boundary effects out of the time interval $I=[0, \frac{N-1}{\fs}]$. Finally, the boundary-effects insensitive representation $\ccF^{\mathrm{ext}}:\RR^{N}\to\CC^{F\times N}$ of $\bx_N$ is given for all $\nu\in\{0,\cdots,F-1\}$, $n\in\{0,\cdots,N-1\}$ by:
\begin{equation}
\ccF^{\mathrm{ext}}(\bx_N)[\nu,n] = \ccF(\tilde \bx_{N+L})[\nu,n]\ .
\label{eq:restriction}
\end{equation}
This amounts to restricting the representation $\ccF(\tilde \bx_{N+L})$ to the current measurement interval of $\bx_N$. For the sake of simplicity, we denote the {\em restriction operator} by $\cR$, where $\cR:\CC^{F\times (N+L)}\to\CC^{F\times N}$. Consequently, we have:
\begin{equation}
\ccF^{\mathrm{ext}}(\bx_N) = \cR\left( \ccF(\tilde \bx_{N+L}) \right) \ .
\label{eq:bound.free.TFR}
\end{equation}
We call $\ccF^{\mathrm{ext}}$ the {\em boundary-free TF representation}.

@@
